Probabilistic Policy Violation Detection with IPFWD

### 2017-02-14 paper/main.tex
Wrote IPFWD section, fix graphs
Reorganize tests, update main.tex

\author{Dr. Phil Nelson and Austin Voecks% <-this % stops a space
\thanks{In collaboration with Dell EMC Isilon}}

@@

IPFWD is a daemon for FreeBSD that updates an early rule in IPFW that has a
chance to accept any packet. The probability of acceptance is updated over time
and dependent on the current system load. To account for this, IPFWD extends
IPFW logs to include the likelihood an undetected policy violation occurred.

This approach is based on the premise that firewall performance can be improved
by reducing the number of rules applied to each packet. Supposing a whitelist
policy, a firewall has to apply every rule to each packet before it's denied.
With IPFWD, we have a chance to accept any packet early and skip any further
computation. Test results shows this reduces the system resources required to
handle the same amount of traffic.

@@

% Firewall basics
Firewalls restrict the types of incoming and outgoing network traffic according
to a security policy. Without a security policy, administrators cannot know
what to allow and disallow. Typical firewalls enforce this policy by blocking
packets that do not match the rules. Whitelist firewalls default action is to
block all traffic, additional rules describe exceptions. Blacklist firewalls
are the opposite in that by default they accept all traffic and additional
rules describe packet types that are not allowed. Since predicting and
blacklisting all possible attacks is infeasible, whitelist firewall rulesets
are the accepted standard and the focus of this paper.

% Why was this firewall chosen?
IPFW is a stateful firewall written for FreeBSD which supports both IPv4 and
IPv6. It is comprised of several components: a kernel firewall filter rule
processor and its integrated packet accounting facility, the logging facility,
NAT, the dummynet traffic shaper, a forward facility, a bridge facility, and
an ipstealth facility.

@@

% First match wins rule system and probabilistic matching
Firewall rules are treated by IPFW in a first-match-wins fashion. IPFW also
contains built-in functionality and rule syntax for probabilistic packet
matching. From the documentation, this feature was intended for load balancing
and other traffic shaping tasks. However it's the core of IPFWD's operation.

@@

% What does it do?
IPFWD increases firewall performance by reducing the average amount of work it
takes to process a packet. Depending on the current system load, IPFWD
increases or decreases the chance the IPFW will immediately accept a given
packet. It does this by adding and maintaining an early rule in the IPFW rule
set of the following form: \verb|prob 0.000 allow ip from any to any|.

@@

In a typical data center environment the vast majority of network traffic is
valid and going to be accepted eventually. However, these rule sets are often
complex and take increasing time to match a given packet depending on the
length. For a rule set of length $n$, we can expect to a packet to encounter
$n/2$ failed matches before matching the correct rule and being accepted.

IPFWD improves the average number of failed matches before successful match
from

@@

% - Log extension
However, while this method increases the speed the system processes packets, it
creates a chance that packets that would have normally been rejected will get
through the firewall. We'll refer to these types of packets as invald.

To account for this, IPFWD extends the IPFW system logs to include information
about the early accpetance probability. From this information, administrators
can extrapolate the probability that additional invalid packets made it through
the firewall for each that is detected. This is discussed further in the
Security Implications section.

% When should this system be used?
IPFWD is not intended to be used on all types of systems. It's application
instead most directly benefits the following types of systems:

@@

% Explain rationale for it acting as it does
IPFWD is based on the assumptions that the majority of network traffic is
valid, malicious events are rare, and

@@

% What does it mean for security if your ignoring some of the packet?

@@

The systems tested all used 64bit FreeBSD 11 Release, with varying CPU type,
CPU speed, RAM and network cards. Isilon and WWU systems were both physical
hardware, while the Virtualized system was in a hypervisor environment hosted
by a third party service.

@@

  Netperf was able to measure CPU utilization on the Virtualized and WWU
  systems, but didn't work on the Isilon systems. Instead, ps was used on those
  systems.  These tests are comparable since both measured total system CPU
  utilization

@@

Firewall decision trees? (Complete Redundancy Detection)

@@

\section{Future Work}
% Firewall rule reordering so we can "fail as early as possible"
Ideally, a whitelist firewall's rule set would be constructed so that the most
commonly applied matches are reached earlier rather than later. IPFWD
accomplishes this task by artificial early stopping. Firewall rule reordering
is a could be another solution to this problem. Given logs of common network
traffic for the system, statistical analysis could show which rules are matched
more often than others. With this information, the rule set could be reordered
so that the most commonly applied rules are earliest.

\section{Conclusion}
% What performance benefit do we get?

@@

\appendices
\section{Additional Figures}
Some text for the appendix.

% use section* for acknowledgement
\section*{Acknowledgment}

The authors would like to thank...

>>>>

\author{Philip Nelson and Austin Voecks% <-this % stops a space
\thanks{Supported in part by Dell EMC Isilon through a grant and hardware}}

@@

IPFWD is a daemon for FreeBSD that updates an early rule in IPFW that has a
chance to accept any packet. The probability of acceptance is updated over time
and dependent on the current system load. In high performance systems this results in an increase in network throughput.

This approach is based on the premise that firewall performance can be improved
by reducing the number of rules applied to each packet. Supposing a whitelist
policy, a firewall has to apply every rule to each packet before it's denied.
With IPFWD, the system has a chance to accept any packet early and skip any further
computation. Test results shows this reduces the system resources required to
handle the same amount of traffic.

@@

% Firewall basics
\subsection{Firewall Basics}
Firewalls restrict the types of incoming and outgoing network traffic according
to a security policy. Without a security policy, administrators cannot know
what to allow and disallow. Typical firewalls enforce the policy by blocking
packets that do not match the rules. The default action for whitelist firewalls is to
block all traffic, additional rules describe exceptions. Blacklist firewalls
are the opposite in that by default they accept all traffic and additional
rules describe packet types that are not allowed. Since predicting and
blacklisting all possible attacks is infeasible, whitelist firewall rulesets
are the accepted standard and the focus of this paper.

% Why was this firewall chosen?
\subsection{Choice of IPFW}
IPFW is a stateful firewall written for FreeBSD which supports both IPv4 and
IPv6. It is comprised of several components: a kernel firewall filter rule
processor and its integrated packet accounting facility, the logging facility,
NAT, the dummynet traffic shaper, a forward facility, a bridge facility, and
an ipstealth facility.

@@

% First match wins rule system and probabilistic matching
Firewall rules are treated by IPFW in a first-match-wins fashion. IPFW also
contains built-in functionality and rule syntax for probabilistic packet
matching. From the documentation, this feature was intended for load balancing
and other traffic shaping tasks. It's the core of IPFWD's operation.

@@

% What does it do?
IPFWD increases firewall performance by reducing the average amount of work it
takes to process a packet. Depending on the current system load, IPFWD
increases or decreases the chance the IPFW will accept a given
packet early. It does this by adding and maintaining an early rule in the IPFW rule
set of the following form: \verb|prob 0.000 allow ip from any to any|.

@@

In a typical data center environment the vast majority of network traffic is
valid and going to be accepted eventually. However, these rule sets are often
complex and take increasing time to match a given packet depending on the
length. For a rule set of length $n$, we can expect to a packet to encounter
$\frac{n}{2}$ failed matches before matching the correct rule and being accepted.

IPFWD improves the average number of failed matches before a successful match
from

@@

% - Log extension
While this method increases the speed the system processes packets, it
creates a chance that packets that would have normally been rejected will get
through the firewall. We'll refer to these types of packets as invald for the purpose of discussion.

In order to provide visibility into this process, IPFWD extends the IPFW system logs to include information
about the early acceptance probability. From this information, administrators
can extrapolate the probability that additional invalid packets made it through
the firewall for each that is detected. This is discussed further in the
\textit{Security Implications} section.

% When should this system be used?
\subsection{Application}
IPFWD is not intended to be used on all types of systems. It's application
instead most directly benefits the following types of systems:

@@

\subsection{Rationale}
% Explain rationale for it acting as it does
IPFWD is based on the assumptions that the majority of network traffic is
valid, malicious events are rare, and TODO

@@

% What does it mean for security if you're ignoring some of the packet?

@@

The systems tested all used x64 FreeBSD 11 Release, with varying CPU type,
CPU speed, RAM and network cards. Isilon and WWU systems were both physical
hardware, while the Virtualized system was in a hypervisor environment hosted
by a third party service.

@@

  Netperf was able to measure CPU utilization on the Virtualized and WWU
  systems, but didn't work on the Isilon systems. Instead, ps was used on those
  systems.  These tests are comparable since both measured total system CPU
  utilization TODO

@@

Firewall decision trees? (Complete Redundancy Detection)
TODO

@@

\section{Future Work}
% Firewall rule reordering so we can "fail as early as possible"
Ideally, a whitelist firewall's rule set would be constructed so that the most
commonly applied matches are reached earlier rather than later. IPFWD
accomplishes this task by artificial early stopping. Firewall rule reordering
is a could be another solution to this problem. Given logs of common network
traffic for the system, statistical analysis could show which rules are matched
more often than others. With this information, the rule set could be reordered
so that the most commonly applied rules are earliest.
TODO

\section{Conclusion}
TODO
% What performance benefit do we get?

@@

\appendices
\section{Additional Figures}
TODO

% use section* for acknowledgement
\section*{Acknowledgment}
TODO
